---
title: "Reproducible Research: Peer Assessment 1"
output: 
  html_document:
    keep_md: true
---
## Requirements for project
```{r,results="hide"}
require(dplyr)
require(ggplot2)
require(lubridate)
```
## Loading and preprocessing the data

```{r}
unzip("activity.zip",files = "activity.csv")
activities <- read.csv("activity.csv",header = TRUE, colClasses = c("numeric","character","numeric"))
activities$date <- as.Date(activities$date)
```

## What is mean total number of steps taken per day?

Histogram of steps per day:
```{r}
groupedActivities <- group_by(activities, date)
stepsPerDay <- summarise(groupedActivities, stepsPerDay = sum(steps,na.rm = TRUE))
qplot(stepsPerDay$stepsPerDay, binwidth=1500)
```

Mean and median of total steps per day:
```{r}
mean(stepsPerDay$stepsPerDay)
median(stepsPerDay$stepsPerDay)
```

## What is the average daily activity pattern?
Timeseries over the intervals, with the average number of steps over days
```{r}
groupedInterval <- group_by(activities,interval)
intervalSummary <- summarise(groupedInterval, averageSteps = mean(steps,na.rm= TRUE))
with(intervalSummary,qplot(x = interval, y = averageSteps, geom = "line"))
```

Maximum number of steps averaged over days

```{r}
intervalSummary[intervalSummary$averageSteps ==max(intervalSummary$averageSteps),1]$interval
```

## Imputing missing values

Total number of NA's:
```{r}
sum(is.na(activities))
```

Show summary of complete table:
```{r}
summary(activities)
```
This shows that only the steps column has missing values.

Imputing missing values with the mean value of that interval over the different days:
```{r}
activitiesImputed <- activities
activitiesImputed$meanSteps <- rep(round(intervalSummary$averageSteps),nrow(activitiesImputed)/nrow(intervalSummary))

activitiesImputed$steps <- ifelse(is.na(activitiesImputed$steps),activitiesImputed$meanSteps,activitiesImputed$steps)
```

Histogram of steps per day on imputed data:
```{r}
groupedActivities <- group_by(activitiesImputed, date)
stepsSummarized <- summarise(groupedActivities, stepsPerDay = sum(steps,na.rm = TRUE))
qplot(stepsSummarized$stepsPerDay, binwidth=1500)
```

Mean and median of total steps per day from imputed data:
```{r}
mean(stepsSummarized$stepsPerDay)
median(stepsSummarized$stepsPerDay)
```

Absolute difference between original data and imputed data for the mean and median of total steps per day:
```{r}
abs(mean(stepsSummarized$stepsPerDay) - mean(stepsPerDay$stepsPerDay))
abs(median(stepsSummarized$stepsPerDay) - median(stepsPerDay$stepsPerDay))
```
There clearly is a difference between the original data and the imputed data. Especially the mean has increased a lot, and also moved towards the median more. This makes sense, since the median is a measure that is less sensitive to noise.

Next to this, you can see that the the histogram has a shape that resembles a normal distribution more than the histogram of the original data.

## Are there differences in activity patterns between weekdays and weekends?
```{r}
factorLevels <- c("weekday","weekend")
dayWeekendFactor <- factor(factorLevels,levels = factorLevels)
activitiesImputed$dayWeekend <- ifelse(wday(activitiesImputed$date) %in% c(1,7),levels(dayWeekendFactor)[2],levels(dayWeekendFactor)[1])
groupedInterval <- group_by(activitiesImputed,interval,dayWeekend)
intervalSummary <- summarise(groupedInterval, averageSteps = mean(steps,na.rm= TRUE))
plot <- ggplot(intervalSummary,aes(x = interval, y = averageSteps)) + geom_line()
plot + facet_grid(dayWeekend ~ .)
```
